Viral Sequence Sorting Tools Evaluation

### 2022-09-09 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Updated VirSorter rule comment.
Playing around with other rule sets for DW data.

Virsorter
    category <= 5: +1
    category == 6: +0.5

@@

## just vibrant
```{r}
viruses$keep_score_vb <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)
```

## vibrant + checkV
```{r}
viruses$keep_score_vb_cv <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)
```

## vibrant + checkV + kaiju
```{r}
viruses$keep_score_vb_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)
```

## vibrant + checkV + kaiju + deepvirfinder
```{r}
viruses$keep_score_vb_cv_kj_dvf <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)
```

## vibrant + virsorter2 + checkV + kaiju
```{r}
viruses$keep_score_vb_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)
```

#all: vibrant + virsorter2 + checkV + kaiju + deepvirfinder
```{r}
viruses$keep_score_all <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T,
                                            include_virsorter = T)
```

## just virsorter2
```{r}
viruses$keep_score_vs2 <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)
```

## virsorter2 + checkV
```{r}
viruses$keep_score_vs2_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)
```

## virsorter2 + checkV + kaiju
```{r}
viruses$keep_score_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)
```

## virsorter2 + checkV + kaiju + deepvirfinder
```{r}
viruses$keep_score_vs2_cv_kj_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)
```

## just checkV
```{r}
viruses$keep_score_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)
```

## checkV + kaiju
```{r}
viruses$keep_score_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)
```

## just kaiju
```{r}
viruses$keep_score_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = T,
                                            include_deepvirfinder = F,
                                            include_virsorter = F)
```

## just deepvirfinder
```{r}
viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = T)
```

## deepvirfinder + checkV
```{r}
viruses$keep_score_dvf_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = T)
```

## deepvirfinder + checkV + kaiju
```{r}
viruses$keep_score_dvf_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T,
                                            include_virsorter = F)
```

## virsorter2 + checkV + kaiju
```{r}
viruses$keep_score_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)
```

## just virsorter
```{r}
viruses$keep_score_vs <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)
```

## virsorter + checkV
```{r}
viruses$keep_score_vs_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)
```

## virsorter + checkV + kaiju
```{r}
viruses$keep_score_vs_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)
```

## virsorter + checkV + kaiju + vibrant + deepvirfinder + virsorter2
```{r}
viruses$keep_score_vs_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T,
                                            include_virsorter = T)
```

## virsorter + checkV + kaiju + vibrant
```{r}
viruses$keep_score_vs_cv_kj_vb <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)
```

## virsorter + checkV + kaiju + virsorter2 + vibrant
```{r}
viruses$keep_score_vs_cv_kj_vb_vs2 <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)
```

## virsorter + checkV + kaiju + virsorter2
```{r}
viruses$keep_score_vs_cv_kj_vs2_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T,
                                            include_virsorter = T)
```

>>>>

Virsorter
    category ==  1,2,4,5: +1
    category == 3,6: +0.5

@@

## vibrant virsorter2 virsorter checkV
```{r}
viruses$keep_score_vbvs2vscv <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_virsorter = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)
```

```{r}
table(viruses$keep_score_vbvs2vscv>=1)
```

```{r}
getting_viral_set_3 <- function(input_seqs,
                                include_vibrant=FALSE, 
                                include_virsorter2=FALSE,
                                include_deepvirfinder=FALSE,
                                include_checkV=FALSE,
                                include_virsorter=FALSE) {

  if (include_vibrant) {
    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- 
      keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 0.5
    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.3
  }

  if (include_virsorter2) {
    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.3
    keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] <- 
      keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] + 0.5
    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
    }

  if (include_virsorter) {
    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.3
    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.3
  }

  if (include_deepvirfinder) {
    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
    keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 1
  }

  if (include_checkV) {
    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0
    keep_score[input_seqs$contig_length>50000 & input_seqs$hallmark==0] <- 0
  }

}
```

## vibrant virsorter2 virsorter checkV
```{r}
viruses$keep_score_vbvs2vscv_3 <- getting_viral_set_3(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_virsorter = T,
                                            include_deepvirfinder = F)
```

```{r}
table(viruses$keep_score_vbvs2vscv_3>=1)
```

```{r}
getting_viral_set_2 <- function(input_seqs,
                                include_vibrant=FALSE, 
                                include_virsorter2=FALSE,
                                include_deepvirfinder=FALSE,
                                include_checkV=FALSE,
                                include_virsorter=FALSE) {

  if (include_vibrant) {
    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- 
      keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 0.5
    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.3
  }

  if (include_virsorter2) {
    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.3
    keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] <- 
      keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] + 0.5
    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
    }

  if (include_virsorter) {
    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.3
    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.3
  }

  if (include_deepvirfinder) {
    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
    keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 1
  }

  if (include_checkV) {
    keep_score[input_seqs$checkv_quality=="low"] <- keep_score[input_seqs$checkv_quality=="low"] + 0.3
    keep_score[input_seqs$checkv_quality=="medium"] <- keep_score[input_seqs$checkv_quality=="medium"] + 0.5
    keep_score[input_seqs$checkv_quality=="high"] <- keep_score[input_seqs$checkv_quality=="medium"] + 1

    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>1] <- 0
  }

}
```

## vibrant virsorter2 virsorter checkV
```{r}
viruses$keep_score_vbvs2vscv_2 <- getting_viral_set_2(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_virsorter = T,
                                            include_deepvirfinder = F)
```

```{r}
table(viruses$keep_score_vbvs2vscv_2>=1)
```
